Weakly Nonlinear MRI: Notes

### 2014-03-11 notes/mri_notes.tex
matrix expansion

>>>>

We now rewrite our system of equations, grouping similar terms into matrices. The system of equations we are rewriting is: \\

$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, J\left(A, \nabla^2 A \right) \, =  \, 2 \partial_z u_{1y} \,+ \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\

$\partial_t u_{1y} \, + \, J\left(\Psi, u_{1y}\right) \, - \, \frac{1}{4\pi} J\left(A, B_{1y}\right) \, + \, \left(2 - \Omega_0 q\right) \partial_z \Psi \, = \, \frac{1}{4\pi}B_0\partial_z B_{1y} \, + \, \frac{1}{\reye} \nabla^2 u_{1y}$ \\

First we have a vector of nonlinear terms:\\

$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - J(A, \nabla^2 A)\\
J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y}) \\ 
-J(A, \Psi) \\
J(\Psi, B_{1y} - J(A, u_{1y}) \\
\end{matrix} \right]$ \\

The time-differential term can be described by $\mathcal{D}\partial_t \mathbf{V}$ where: \\

$\mathcal{D} = \left[\begin{matrix}
\nabla^2 & 0 & 0 & 0 \\
0 & 1& 0 & 0 \\
0 & 0 & 1 & 0\\
0 & 0 & 0 & 1 \\
\end{matrix}\right]$ \\

The righthand side of the equations can be split into three matrices: one containing the Reynolds numbers, one containing the magnetic field strength, and one containing the shear terms. That is: \\

$RHS = \left[\begin{matrix}
\frac{1}{\reye} \nabla^4 & 0 & 0 & 0 \\
0 & \frac{1}{\reye}\nabla^2 & 0 & 0 \\
0 & 0 & \frac{1}{\reym}\nabla^2 & 0 \\
0 & 0 & 0 & \frac{1}{\reym}\nabla^2 \\ 
\end{matrix}\right] + \left[\begin{matrix}
0 & 0 & \frac{1}{4\pi}B_0 \partial_z \nabla^2 & 0 \\
0 & 0 & 0 & \frac{1}{4\pi} B_0 \partial_z \\
B_0 \partial_z & 0 & 0 & 0 \\
0 & B_0 \partial_z & 0 & 0 \\ \end{matrix}\right]$ \\

$+ \left[\begin{matrix}
0 & 2\partial_z & 0 & 0 \\
\left(\Omega_0 - 2\right)q\partial_z & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & -q\Omega_0\partial_z & 0 \\ \end{matrix}\right]$ \\

We will now regroup matrices by $\mathcal{O}(\partial_z)$. The first matrix is rewritten as three matrices, keeping in mind that $\nabla^4 = \partial_x^4 + \partial_z^4 + 2\partial_x\partial_z^2$: \\

$\left[\begin{matrix}
\frac{1}{\reye} \nabla^4 & 0 & 0 & 0 \\
0 & \frac{1}{\reye}\nabla^2 & 0 & 0 \\
0 & 0 & \frac{1}{\reym}\nabla^2 & 0 \\
0 & 0 & 0 & \frac{1}{\reym}\nabla^2 \\ 
\end{matrix}\right] = \left[\begin{matrix}
\frac{1}{\reye}\partial_x^4 & 0 & 0 & 0 \\
0 & \frac{1}{\reye}\partial_x^2 & 0 &0 \\
0 & 0 & \frac{1}{\reym}\partial_x^2 & 0 \\
0 & 0 & 0 & \frac{1}{\reym}\partial_x^2 \\ \end{matrix}\right] $ \\

$+ \left[\begin{matrix}
\frac{1}{\reye}\partial_z^4 & 0 & 0 & 0 \\
0 & \frac{1}{\reye}\partial_z^2 & 0 & 0 \\
0 & 0 & \frac{1}{\reym} \partial_z^2 & 0 \\
0 & 0 & 0 & \frac{1}{\reym} \partial_z^2 \\ \end{matrix} \right] + \left[ \begin{matrix}
\frac{1}{\reye} 2\partial_x^2\partial_z^2 & 0 & 0 & 0 \\
0 & 1 & 0 & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 0 & 1\\ \end{matrix}\right]$

The second matrix involves $B_0$, which we rewrite in terms of the tuning parameter $\epsilon^2 \equiv 1 - B_0$.
